curatedMetagenomicData

### 2019-12-03 combodata_NS.Rmd
updated

```{r setup, include=FALSE}

#BiocManager::install("curatedMetagenomicData")
#BiocManager::install("ExperimentHub")
library(ExperimentHub)
library(curatedMetagenomicData)
suppressPackageStartupMessages(library(ExperimentHub))
library(phyloseq)
library(microbiome)
library(forcats)
library(stargazer)
library(ggpubr)
library(tidyverse)
library(cowplot)
library(vegan)
library(broom)
library(tsnemicrobiota)
library(latex2exp)
library(wesanderson)

@@

# the relab=FALSE is essential as it Absolute Raw Count Data
pseq = ExpressionSet2phyloseq(eset,relab=FALSE ,simplify = TRUE, phylogenetictree = TRUE)
orig_pseq <- pseq
#For the MetaPhlAn2 bugs datasets (but not other data types), you gain a lot of taxonomy-aware, ecological analysis and plotting by conversion to a phyloseq class object. curatedMetagenomicData provides the ExpressionSet2phyloseq() function to make this easy:
data(atlas1006)
atlas.pseq <- baseline(atlas1006)

@@

## R Markdown

```{r}
# Create mnd variable
pseq <- orig_pseq
# grabs the nationality from phyloseq
mnd.country <- get_variable(pseq, "country" )
table(mnd.country)

@@

# Cut off is 1.8  Low

mnd.country <- fct_collapse(mnd.country, LOW = c("DEU","ESP","ITA","AUT","DNK","LUX" ) , HIGH = c("NOR","FIN","FRA","NLD","GBR","USA","CAN","SWE") )

table(mnd.country)
#mnd.country <- fct_collapse(mnd.country, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") ) "ITA","AUT","DEU","ESP","FRA","NLD"

# reorder
mnd.country <- factor(mnd.country, levels = (c("LOW", "HIGH")))
levels(mnd.country)

# creates a new variable in the phyloseq called mnd
sample_data(pseq)$mnd = mnd.country

# checks that it has worked.
get_variable(pseq, "mnd")

# THIS IS FOR THE ATLAS 1006 
# Create mnd variable
# grabs the nationality from phyloseq
mnd.nationality <- get_variable(atlas.pseq, "nationality" )

# this collapses the two varialbes into a new variable AB in this case. The command for multiple changes is fct_collapse(x, AB = c("A","B"), DE = c("D","E"))
mnd.nationality <- fct_collapse(mnd.nationality, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") )

# reorder
mnd.nationality <- factor(mnd.nationality, levels = (c("LOW", "MEDIUM", "HIGH")))
levels(mnd.nationality)

# creates a new variable in the phyloseq called mnd
sample_data(atlas.pseq)$mnd = mnd.nationality

# checks that it has worked.
get_variable(atlas.pseq, "mnd")

@@

Defining the functions 
```{r define the functions}

@@

# THis is required
mylab <- function(v1) {
  deparse(substitute(v1))
}
# aplha_d runs the diversity for each country or  groupoing and then compares the three groups
# example is alpha_d(pseq.subset,"TEST_3") this will generate a png file called TEST_3.png in the 
# currect working directory. This can be checked with getwd()
alpha_d_two <- function(pseq.subset,title_text) {
   # gets the meta data 
   pseq.meta <- meta(pseq.subset)
   # generates the diversites
   tab <- microbiome::alpha(pseq.subset, index = "shannon")

   # creates Shannon
   pseq.meta$Shannon <- tab$diversity_shannon
   mnd <- levels(pseq.meta$mnd) # get the variables
  # Make a pairwise list that we want to compare.
  mnd.pairs <- combn(seq_along(mnd), 2, simplify = FALSE, FUN = function(i)mnd[i])
  mnd.pairs <- list( c("HIGH","LOW"))

   # Compare differences in Shannon index between mnd group of the study subjects
  pseq.meta$mnd <- factor(pseq.meta$mnd, levels = c("LOW","HIGH"),  exclude = "NA")
  # on to the graphs!!! country 
  p1<-   ggviolin(pseq.meta, x = "nationality", y = "Shannon",
 add = "boxplot", fill = "nationality", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")
# the number in the palettes must equal the number of groups in bmi.pairs

@@

    p2 <<- p2 + stat_compare_means(comparisons = mnd.pairs) + xlab("MND Prevalence") + theme(legend.position="none")
    #this just generates the titles for the output
    title <- ggdraw() + 
    draw_label(
    title_text,
    fontface = 'bold',
    x = 0,
    hjust = 0
    ) +
    theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
    )

    p3 <<-plot_grid(title,p1,p2, align = "v" , ncol = 1, labels = c('','A', 'B'),rel_heights = c(0.1, 1,1.2))
    ggsave(paste("Alpha_diversity_",title_text,".png", sep = "" ), device = "png")
}

# THis includes the analysis as well so can take some time to run. the graphs are generated with the label of beta diversity. It requires an input of a title. 
beta_d_two <- function(pseq.subset,title_text) {
  #NMDS ordination on Bray-Curtis distance
  p <<-  boxplot(list(Low = p.low.d, High = p.high.d))
 # p1 = plot_ordination(pseq.subset, pseq.nmds, type = "taxa", color = "mnd", title = "Taxa by Phylum") # I dont think that we need this any more. 
  #p1 + facet_wrap(~Phylum, 3)
  pseq.nmds <- ordinate(pseq.subset, "NMDS", "bray")
  p2 <- plot_ordination(pseq.subset, pseq.nmds, type = "sample", color = "mnd")

@@

pre_perma_d_two <- function(pseq.subset) { 
  p.low <<- subset_samples(pseq.subset, mnd == "LOW")
  p.low.d <<- divergence(p.low)
  p.high <<- subset_samples(pseq.subset, mnd == "HIGH")
  p.high.d <<- divergence(p.high)
  p <<- boxplot(list(Low = p.low.d, High = p.high.d))

@@

# THis calcualtes the group PERMANOVA perma_d(p.low,p.high,"TESTHigh_vesus_Low")
perma_d_two <- function(title_text) { # x & y must be either p.low,p.med or p.high
  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
  pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 20)  + annotate("text", x = 2.5, y = 0.02, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa")  
  ggsave(paste("PERMANOVA_",title_text,".png", sep = "" ), device = "png",width = 210, height = 297, units = "mm")
  ggsave("test_permanova.png", device = "png",width = 297, height = 210, units = "mm")
 }

testpseq <- atlas1006 %>%
   subset_samples(DNA_extraction_method == "r") %>%
   aggregate_taxa(level = "Phylum") %>%
   transform(transform = "compositional")
p <- plot_composition(testpseq, sample.sort = "Firmicutes",
         otu.sort = "abundance", verbose = TRUE) +
     scale_fill_manual(values = default_colors("Phylum")[taxa(testpseq)])

comp_d <- function(pseq,subset,title_text) {
  png(paste(title_text,"Phylum_abundance.png",sep="_"))
  p <- plot_bar(pseq.subset,"Phylum") + theme(axis.text.x = element_text(face="italic"))
  print(p)

  prev.otu <- plot_taxa_prevalence(pseq.subset, "Phylum" , detection = 5)
  prev.otu + theme(legend.position="none")
  ggsave(paste(title_text,"phylum_prevalence.png",sep="_"))
  tab <- plot_taxa_prevalence(pseq.subset, "Phylum", detection = 5)
  tab + theme(legend.position="none")
  ggsave(paste(title_text,"Phylum_prevalence_detection.png",sep="_"))

  pseq.ph <- aggregate_taxa(pseq.subset, "Phylum")
  pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
  plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample") +
    theme(legend.position = "bottom") + theme_bw() +
    theme(axis.text.x=element_blank()) +
    c + scale_fill_brewer(palette = "Paired") + guides(fill=guide_legend("Phylum"))
  print(plot.comp.rel)
  ggsave("phylum_relative_abundance11.png")
  plot.comp.rel + theme(legend.position="none")
  ggsave("phylum_relative_abundance12.png")

  pseq.ph <- aggregate_taxa(pseq.subset, "Phylum")
  pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
  plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND   Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))
  plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
  ggsave("phylum_abundance_bar.png")
  # Top 5
  top5P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:5]

  top5P = subset_taxa(pseq, Genus %in% names(top5P.names))
u  pseq.ph.rel <- microbiome::transform(top5P, "compositional")
  plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))

  plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma")

  # Top 10
  top10P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:10]

  top10P = subset_taxa(pseq, Genus %in% names(top10P.names))
  pseq.ph.rel <- microbiome::transform(top10P, "compositional")

  plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))

  plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
  # new
  p.new <- aggregate_taxa(pseq, "Phylum")
  p.new.rel <- microbiome::transform(p.new, "compositional")

  plot.p.new.rel <- plot_composition(p.new.rel, average_by = "mnd", x.label = " ", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text( face="italic"))

  plot.p.new.rel + scale_fill_brewer(palette = "Paired") 
  ggsave("pnew_abundance_bar.png")

  # Top Phylum Heatmap
  pseq.ph <- aggregate_taxa(pseq, "Phylum")

  pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")

  plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample", plot.type = "heatmap")

  plot(plot.comp.rel+ coord_flip()) + theme(axis.text.x=element_blank())
  ggsave("phylum_histogram.png")

  #p.new heatmap
  pseq.n.ph <- aggregate_taxa(pseq, "Phylum")
  pseq.n.ph.rel <- microbiome::transform(pseq.n.ph, "compositional")

  plot.n.comp.rel <- plot_composition(pseq.n.ph.rel, x.label = "sample", plot.type = "heatmap")

  p <- plot.n.comp.rel + coord_flip()
  p + theme(axis.text.x=element_blank()) + labs(y = "Sample")
  ggsave("pnew_histogram.png")

  # Newwork
  png("network_bray_phylum.png")
  jg = make_network(pseq, "taxa", "bray", 0.3)
  plot_network(jg, pseq, "taxa", color = "Phylum", line_weight = 0.4, label = NULL)
  dev.off()
  #Community-composition-plots
  # Create a data table for ggplot
# agglomerate at phylum level
pseq_phy <- tax_glom(pseq.subset, taxrank = "Phylum")

# Transform to rel. abundance (or use ps0.ra)
pseq_phyl <- transform_sample_counts(pseq_phy, function(x) {x/sum(x)})

# Melt to long format for easy ggploting
pseq_phylum <- psmelt(pseq_phyl)

     # filter(Abundance > 0.01)                             # Filter out low abundance taxa
# Plot - Phylum
    p.ra.phylum <- ggplot(pseq_phylum, aes(x = sample, y = Abundance, fill = Phylum)) + 
    geom_bar(stat = "identity", width = 1) +
    facet_wrap(~mnd, scales = "free_x", nrow = 4, ncol = 7) +
    theme(axis.text.x = element_blank()) +
    theme(axis.title.x = element_blank()) +
    labs(title = "Abundant Phylum (> 1%)")

  p <- p.ra.phylum + scale_fill_viridis(discrete = TRUE, option = "A")
  plot(p)
  ggsave("phylum_abun_mnd.png")

 # these are other functions required along the way.

    # transform counts to n
  physeq.scale <- transform_sample_counts(physeq, 
    function(x) {(n * x/sum(x))}
  )

  # Pick the rounding functions
  if (round == "floor"){
    otu_table(physeq.scale) <- floor(otu_table(physeq.scale))
  } else if (round == "round"){
    otu_table(physeq.scale) <- myround(otu_table(physeq.scale))
  }

  # Prune taxa and return new phyloseq object
  physeq.scale <- prune_taxa(taxa_sums(physeq.scale) > 0, physeq.scale)
  return(physeq.scale)
}

@@

## Check OTUs/Phylums
Prune(remove) taxa because we want to remove OTUs/taxa that aren't in any of the samples
Don't need to prune this taxa because OTUs are found in all samples.
Check this everytime you use the subset_samples function because of the pseq object being filtered. 
```{r check OTUs}
# Check if any OTUs are not present in any samples
any(taxa_sums(pseq.subset) == 0)
# FALSE, therefore, OTUs are found in all samples and we don't need to prune the taxa. If the answer is TRUE, some OTUs are not present in any samples. This is usually the case when data is subset to remove some samples. OTUs unique to those sample are not removed along with the samples. Therefore, it is important to check this everytime the phyloseq object is filtered for samples using `subset_samples` function.

@@

## Prepare Data for Alpha Diversity Visualisation
```{r prepare data for alpha diversity visualisation}

@@

# Remove   everything without a country

@@

pseq.subset <- subset_samples(pseq.c, !is.na(mnd) & disease == "healthy" & age_category != "child" & age_category != "newborn" & age_category != "schoolage" & antibiotics_current_use != "yes")

>>>>

## Setup
```{r setup, include=FALSE}

#BiocManager::install("curatedMetagenomicData")
#BiocManager::install("ExperimentHub")
library(ExperimentHub)
library(curatedMetagenomicData)
suppressPackageStartupMessages(library(ExperimentHub))
library(phyloseq)
library(microbiome)
library(forcats)
library(stargazer)
library(ggpubr)
library(tidyverse)
library(cowplot)
library(vegan)
library(broom)
library(tsnemicrobiota)
library(latex2exp)
library(wesanderson)
library(table1)

@@

# the relab=FALSE is essential as it Absolute Raw Count Data
pseq = ExpressionSet2phyloseq(eset,relab=FALSE ,simplify = TRUE, phylogenetictree = TRUE)
orig_pseq <- pseq

#For the MetaPhlAn2 bugs datasets (but not other data types), you gain a lot of taxonomy-aware, ecological analysis and plotting by conversion to a phyloseq class object. curatedMetagenomicData provides the ExpressionSet2phyloseq() function to make this easy:
data(atlas1006)
atlas.pseq <- baseline(atlas1006) # only uses the baseline samples. This is n=1006

@@

## MND Countries
```{r tables}

# THIS FOR THE CURATED DATA SET

# creates a grouping varaible for BMI
n <- length(sample_data(pseq)$BMI)
data <- data.frame(id=1:n)
data$bmi <- sample_data(pseq)$BMI # creates a new variable with BMI in it
# Classification based upon the ATlas data
data$bmi[data$bmi <18.5] <- "underweight"
data$bmi[data$bmi >=45] <- "super obese"
data$bmi[data$bmi >= 18.5 & data$bmi <25] <- "lean"
data$bmi[data$bmi >= 25 & data$bmi <30] <- "overweight"
data$bmi[data$bmi >= 30 & data$bmi <35] <- "obese"
data$bmi[data$bmi >= 35 & data$bmi <40] <- "severe obese"
data$bmi[data$bmi >= 40 & data$bmi <45] <- "morbid obese"
data$bmi <- factor(data$bmi)

# creates a new variable in the phyloseq called mnd
sample_data(pseq)$bmi_group = data$bmi
# Check that it works
get_variable(pseq, "bmi_group")

# Create mnd variable
#pseq <- orig_pseq
# grabs the nationality from phyloseq
c.mnd.country <- get_variable(pseq, "country" )
table(c.mnd.country)

@@

#Adjsuted to match exactly the ATLAS data set

#central Europe = Belgium, Demark, Netherlands, Germany
#southern Europe = France, Italy, Serbia, Spain
#UKIE
#Scandinavia = Norway, Sweden Finland
#Eastern Europe = Poland

c.mnd.atlas <- fct_collapse(c.mnd.country, Scandinavia = c("NOR","SWE","FIN") ,CentralEurope = c("DNK","DEU","NLD"), SouthEurope = c("ITA","ESP","FRA"), US = "USA", UKIE = "GBR" )

#mnd.country <- fct_collapse(mnd.country, LOW = c("NOR","SWE","FIN","DNK","LUX") , MEDIUM = c("ITA","AUT","DEU","ESP","FRA","NLD"), HIGH = c("GBR","USA","CAN") )

#mnd.country <- fct_collapse(mnd.country, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") )

# reorder
# NOTE THERE IS NO EASTERN EUROPE
c.mnd.nationality <- fct_collapse(c.mnd.atlas, LOW = c("Scandinavia") , MEDIUM = c("SouthEurope","CentralEurope" ), HIGH = c("US","UKIE") )

c.mnd.nationality <- factor(c.mnd.nationality, levels = (c("LOW", "MEDIUM", "HIGH")))
levels(c.mnd.nationality)

# creates a new variable in the phyloseq called mnd
sample_data(pseq)$mnd = c.mnd.nationality

# checks that it has worked.
get_variable(pseq, "mnd")
table(get_variable(pseq, "mnd"))

###############################################

# THIS IS FOR THE ATLAS 1006 
# Create mnd variable
# grabs the nationality from phyloseq
a.mnd.nationality <- get_variable(atlas.pseq, "nationality" )

# this collapses the two varialbes into a new variable AB in this case. The command for multiple changes is fct_collapse(x, AB = c("A","B"), DE = c("D","E"))
a.mnd.nationality <- fct_collapse(a.mnd.nationality, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope" ), HIGH = c("US","UKIE") )

# reorder
a.mnd.nationality <- factor(a.mnd.nationality, levels = (c("LOW", "MEDIUM", "HIGH")))
levels(a.mnd.nationality)

# creates a new variable in the phyloseq called mnd
sample_data(atlas.pseq)$mnd = a.mnd.nationality
#sample_data(atlas.pseq)$country = sample_data(atlas.pseq)$nationality
# checks that it has worked.
table(get_variable(atlas.pseq, "mnd"))

@@

##Functions 
```{r function}

@@

my_cur_table <- function(pseq.subset){
    n <- length(sample_data(pseq.subset)$mnd)
    data <- data.frame(id=1:n)
    data$mnd <- sample_data(pseq.subset)$mnd
    data$bmi <- sample_data(pseq.subset)$BMI
    data$bmi_group <- sample_data(pseq.subset)$bmi_group
    data$age <- sample_data(pseq.subset)$age
    data$antibiotics_current_use <- sample_data(pseq.subset)$antibiotics_current_use
    data$smoker <- sample_data(pseq.subset)$smoker
    data$pregnant <- sample_data(pseq.subset)$pregnant
    data$alcohol <- sample_data(pseq.subset)$alcohol
    data$age_category <- sample_data(pseq.subset)$age_category
    table1(~ age  +  bmi + bmi_group + smoker + alcohol + age_category | mnd, data=data)
}

my_atlas_table <- function(atlas.pseq){
    n <- length(sample_data(atlas.pseq)$mnd)
    data <- data.frame(id=1:n)
    data$mnd <- sample_data(atlas.pseq)$mnd
    data$bmi <- sample_data(atlas.pseq)$bmi_group
    data$age <- sample_data(atlas.pseq)$age
    table1(~ age  +  bmi | mnd, data=data)
}

# THis is required
mylab <- function(v1) {
  deparse(substitute(v1))
}
# aplha_d runs the diversity for each country or  groupoing and then compares the three groups
# example is alpha_d(pseq.subset,"TEST_3") this will generate a png file called TEST_3.png in the 
# currect working directory. This can be checked with getwd()
alpha_d <- function(pseq.subset,title_text) {
  # gets the meta data 
  pseq.meta <- meta(pseq.subset)
  # generates the diversites
  tab <- microbiome::alpha(pseq.subset, index = "shannon")

  # creates Shannon
  pseq.meta$Shannon <- tab$diversity_shannon
  mnd <- levels(pseq.meta$mnd) # get the variables
  # Make a pairwise list that we want to compare.
  mnd.pairs <- combn(seq_along(mnd), 2, simplify = FALSE, FUN = function(i)mnd[i])
  mnd.pairs <- list( c("LOW","MEDIUM"), c("HIGH","LOW"), c("MEDIUM","HIGH"))

  # Compare differences in Shannon index between mnd group of the study subjects
  pseq.meta$mnd <- factor(pseq.meta$mnd, levels = c("LOW","MEDIUM","HIGH"),  exclude = "NA")
  # on to the graphs!!!
  print("The basic aloha diversity analysis has been completed")
  # This country versus nationality
  p1<-   ggviolin(pseq.meta, x = "country", y = "Shannon",
                  add = "boxplot", fill = "country", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")
  print("First graph has completed")
  # the number in the palettes must equal the number of groups in bmi.pairs

@@

  p2 <<- p2 + stat_compare_means(comparisons = mnd.pairs) + xlab("MND Prevalence") + theme(legend.position="none")

  print("Second graph has completed")
  #this just generates the titles for the output
  title <- ggdraw() + 
    draw_label(
      title_text,
      fontface = 'bold',
      x = 0,
      hjust = 0
    ) +
    theme(
      # add margin on the left of the drawing canvas,
      # so title is aligned with left edge of first plot
      plot.margin = margin(0, 0, 0, 7)
    )

  p3 <<-plot_grid(title,p1,p2, align = "v" , ncol = 1, labels = c('','A', 'B'),rel_heights = c(0.1, 1,1.2))

  print("Third graph has completed")
  ggsave(paste("/Users/Nik/Alpha_diversity_",title_text,".png", sep = "" ), device = "png")
  #ggsave("/Users/Nik/Alpha_diversity_curate.png")
}

# THis includes the analysis as well so can take some time to run. the graphs are generated with the label of beta diversity. It requires an input of a title. 
beta_d <- function(pseq.subset,title_text) {
  #NMDS ordination on Bray-Curtis distance
  p <<-  boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))
 # p1 = plot_ordination(pseq.subset, pseq.nmds, type = "taxa", color = "mnd", title = "Taxa by Phylum") # I dont think that we need this any more. 
  #p1 + facet_wrap(~Phylum, 3)
  pseq.nmds <- ordinate(pseq.subset, "NMDS", "bray")
  p2 <- plot_ordination(pseq.subset, pseq.nmds, type = "sample", color = "mnd")

@@

pre_perma_d <- function(pseq.subset) { 
  p.low <<- subset_samples(pseq.subset, mnd == "LOW")
  p.low.d <<- divergence(p.low)
  p.med <<- subset_samples(pseq.subset, mnd == "MEDIUM")
  p.med.d <<- divergence(p.med)
  p.high <<- subset_samples(pseq.subset, mnd == "HIGH")
  p.high.d <<- divergence(p.high)
  p <<- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))

@@

# THis calcualtes the group PERMANOVA perma_d(p.low,p.high,"TESTHigh_vesus_Low")
perma_d <- function(x,y,title_text) { # x & y must be either p.low,p.med or p.high
  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
  pseqLvH <- merge_phyloseq(x,y) # pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 20)  + annotate("text", x = 2.5, y = 0.02, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa")  
  ggsave(paste("PERMANOVA_",title_text,".png", sep = "" ), device = "png",width = 210, height = 297, units = "mm")
  #ggsave("test_permanova.png", device = "png",width = 297, height = 210, units = "mm")
 }

# testpseq <- atlas1006 %>%
#    subset_samples(DNA_extraction_method == "r") %>%
#    aggregate_taxa(level = "Phylum") %>%
#    transform(transform = "compositional")
# p <- plot_composition(testpseq, sample.sort = "Firmicutes",
#          otu.sort = "abundance", verbose = TRUE) +
#      scale_fill_manual(values = default_colors("Phylum")[taxa(testpseq)]) 
# 
# comp_d <- function(pseq,subset,title_text) {
#   png(paste(title_text,"Phylum_abundance.png",sep="_"))
#   p <- plot_bar(pseq.subset,"Phylum") + theme(axis.text.x = element_text(face="italic"))
#   print(p)
# 
#   prev.otu <- plot_taxa_prevalence(pseq.subset, "Phylum" , detection = 5)
#   prev.otu + theme(legend.position="none")
#   ggsave(paste(title_text,"phylum_prevalence.png",sep="_"))
#   tab <- plot_taxa_prevalence(pseq.subset, "Phylum", detection = 5)
#   tab + theme(legend.position="none")
#   ggsave(paste(title_text,"Phylum_prevalence_detection.png",sep="_"))
#   
#   pseq.ph <- aggregate_taxa(pseq.subset, "Phylum")
#   pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
#   plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample") +
#     theme(legend.position = "bottom") + theme_bw() +
#     theme(axis.text.x=element_blank()) +
#     c + scale_fill_brewer(palette = "Paired") + guides(fill=guide_legend("Phylum"))
#   print(plot.comp.rel)
#   ggsave("phylum_relative_abundance11.png")
#   plot.comp.rel + theme(legend.position="none")
#   ggsave("phylum_relative_abundance12.png") 
#   
#   pseq.ph <- aggregate_taxa(pseq.subset, "Phylum")
#   pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
#   plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND   Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))
#   plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
#   ggsave("phylum_abundance_bar.png")
#   # Top 5
#   top5P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:5]
# 
#   top5P = subset_taxa(pseq, Genus %in% names(top5P.names))
# u  pseq.ph.rel <- microbiome::transform(top5P, "compositional")
#   plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))
# 
#   plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
#   
#   # Top 10
#   top10P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:10]
# 
#   top10P = subset_taxa(pseq, Genus %in% names(top10P.names))
#   pseq.ph.rel <- microbiome::transform(top10P, "compositional")
# 
#   plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))
# 
#   plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
#   # new
#   p.new <- aggregate_taxa(pseq, "Phylum")
#   p.new.rel <- microbiome::transform(p.new, "compositional")
# 
#   plot.p.new.rel <- plot_composition(p.new.rel, average_by = "mnd", x.label = " ", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text( face="italic"))
# 
#   plot.p.new.rel + scale_fill_brewer(palette = "Paired") 
#   ggsave("pnew_abundance_bar.png")
#   
#   # Top Phylum Heatmap
#   pseq.ph <- aggregate_taxa(pseq, "Phylum")
# 
#   pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
# 
#   plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample", plot.type = "heatmap")
# 
#   plot(plot.comp.rel+ coord_flip()) + theme(axis.text.x=element_blank())
#   ggsave("phylum_histogram.png")
#   
#   #p.new heatmap
#   pseq.n.ph <- aggregate_taxa(pseq, "Phylum")
#   pseq.n.ph.rel <- microbiome::transform(pseq.n.ph, "compositional")
# 
#   plot.n.comp.rel <- plot_composition(pseq.n.ph.rel, x.label = "sample", plot.type = "heatmap")
# 
#   p <- plot.n.comp.rel + coord_flip()
#   p + theme(axis.text.x=element_blank()) + labs(y = "Sample")
#   ggsave("pnew_histogram.png")
#   
#   # Newwork
#   png("network_bray_phylum.png")
#   jg = make_network(pseq, "taxa", "bray", 0.3)
#   plot_network(jg, pseq, "taxa", color = "Phylum", line_weight = 0.4, label = NULL)
#   dev.off()
#   #Community-composition-plots
#   # Create a data table for ggplot
# # agglomerate at phylum level
# pseq_phy <- tax_glom(pseq.subset, taxrank = "Phylum")                  
# 
# # Transform to rel. abundance (or use ps0.ra)
# pseq_phyl <- transform_sample_counts(pseq_phy, function(x) {x/sum(x)}) 
# 
# # Melt to long format for easy ggploting
# pseq_phylum <- psmelt(pseq_phyl) 
#  
#      # filter(Abundance > 0.01)                             # Filter out low abundance taxa
# # Plot - Phylum
#     p.ra.phylum <- ggplot(pseq_phylum, aes(x = sample, y = Abundance, fill = Phylum)) + 
#     geom_bar(stat = "identity", width = 1) +
#     facet_wrap(~mnd, scales = "free_x", nrow = 4, ncol = 7) +
#     theme(axis.text.x = element_blank()) +
#     theme(axis.title.x = element_blank()) +
#     labs(title = "Abundant Phylum (> 1%)")
# 
#   p <- p.ra.phylum + scale_fill_viridis(discrete = TRUE, option = "A")
#   plot(p)
#   ggsave("phylum_abun_mnd.png")
# 
#  # these are other functions required along the way. 
# 
#     # transform counts to n
#   physeq.scale <- transform_sample_counts(physeq, 
#     function(x) {(n * x/sum(x))}
#   )
#   
#   # Pick the rounding functions
#   if (round == "floor"){
#     otu_table(physeq.scale) <- floor(otu_table(physeq.scale))
#   } else if (round == "round"){
#     otu_table(physeq.scale) <- myround(otu_table(physeq.scale))
#   }
#   
#   # Prune taxa and return new phyloseq object
#   physeq.scale <- prune_taxa(taxa_sums(physeq.scale) > 0, physeq.scale)
#   return(physeq.scale)
# }
#

@@

```{r check OTUs}
# Check if any OTUs are not present in any samples
any(taxa_sums(pseq.subset) == 0)
# FALSE, therefore, OTUs are found in all samples and we don't need to prune the taxa. If the answer is TRUE, some OTUs are not present in any samples. This is usually the case when data is subset to remove some samples. OTUs unique to those sample are not removed along with the samples. Therefore, it is important to check this everytime the phyloseq object is filtered for samples using `subset_samples` function.

@@

## Filtering
```{r  data filtering}

@@

# Rmove   everything without a country

@@

#pseq.subset <- subset_samples(pseq.c, !is.na(mnd) & disease == "healthy" & age_category != "child" & age_category != "newborn" & age_category != "schoolage" & antibiotics_current_use != "yes")
